Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-20 WSSSPE1_report/WSSSPE1_summary_report.tex
more text to sec 9 subsections

\subsection{Utility Software}
Utility software potentially has the broadest community impact of the three
categories. This is especially relevant in the modern era of social computing. In the context of
this paper, a utility software project could be defined as general purpose software
applicable to one or many generic tasks and/or enabling other software to run
by providing a suitable environment. Examples are collaborative development
framework such as \emph{GitHub} and \emph{Bitbucket}, workflow and scripting
frameworks such as \emph{Galaxy}, \emph{Swift}, and \emph{VisTrails},
and visualization frameworks such as \emph{visIT}. Often projects start with
specific, small set of requirements and users in mind but expand to provide
more generic functionality and appeal to a wider user community. Owing to its generic nature, utility software finds a wider users communities which in turn leads to a higher participation in development. Consequently, the process becomes user-driven and self sustaining. Software developed in such scenario often benefits from community driven development, deployment and maintenance best practices.

For instance the \emph{Galaxy} project has significantly benefitted from user community participation in its development. It follows agile software development practices and implements standard practices such as test-driven development and socialized bug managing practices via \emph{trello}. Galaxy \emph{histories} and \emph{toolshed} offers easy community sharing of data and tools further promoting a collaborative environment.

\subsection{Scientific Software}
Special purpose software that was developed specifically for a target use-case
scenario or keeping some fixed/frozen requirements in mind.  Domain science
oriented software developed to solve domain science problems, e.g., MINRES-QLP.

\subsection{Infrastructure Software}
Infrastructure oriented software developed to better utilize new
infrastructure, e.g., MVAPICH.
%
%\begin{itemize}
%  \item mvapich, Cactus, DUNE
%  \item gateways, github, VisIT, visTrails, PAAS
%  \item SimGrid, Swift, Galaxy
%  \item FastMath, R/qtl, MINRES-QLP
%  \item OWL/RDF, AMR/multiphysics, Globus-Genomics
%\end{itemize}
%

>>>>

\subsection{Utility Software}
Utility software potentially has the broadest community impact of the three
categories. This is especially relevant in the modern era of social computing. In the context of
this paper, a utility software project could be defined as general purpose software
applicable to one or many generic tasks and/or enabling other software to run
by providing a suitable environment. Examples are collaborative development
framework such as \emph{GitHub} and \emph{Bitbucket}, workflow and scripting
frameworks such as \emph{Galaxy}, \emph{Swift}, \emph{Globus} and \emph{VisTrails},
and visualization frameworks such as \emph{visIT}. Often projects start with
specific, small set of requirements and users in mind but expand to provide
more generic functionality and appeal to a wider user community. Owing to its generic nature, utility software finds a wider users communities which in turn leads to a higher participation in development. Consequently, the process becomes user-driven and self sustaining. Software developed in such scenario often benefits from community driven development, deployment and maintenance best practices. For instance the \emph{Galaxy} project has significantly benefitted from user community participation in its development. It follows agile software development practices and implements standard practices such as test-driven development and socialized bug managing practices via \emph{trello}. Galaxy \emph{histories} and \emph{toolshed} offers easy community sharing of data and tools further promoting a collaborative environment.

Another paradigm in utility software is the software delivered as service over the web.

\subsection{Scientific Software}
Special purpose software that was developed specifically for a target use-case
scenario or keeping some fixed/frozen requirements in mind. Software projects pertaining to specific scientific domain often tend to be in a niche and the user community tends to be small to medium. Accuracy and optimization are some of the paramount requirements of most scientific software. Advantages of a smaller codebase and requirements include stability, ease of installation, configurability and integration into larger systems. While the software can stay stable and require relatively low maintenance, the responsibility is often on the shoulders of a very few developers. Community participation in the development tends to be low. Development is linear and simplistic with a limited scope to follow software best practices. Domain science oriented software developed to solve  specific problems, e.g., \emph{MINRES-QLP}, \emph{FASTMath}.

\subsection{Infrastructure Software}
Infrastructure oriented software developed to better utilize new
infrastructure, e.g., MVAPICH. Explores uncharted territories. Highest risk. Long requirement-development-test cycles. Wide community support but tends to be handled by specialists rather then end users.

%
%\begin{itemize}
%  \item mvapich, Cactus, DUNE
%  \item gateways, github, VisIT, visTrails, PAAS
%  \item SimGrid, Swift, Galaxy
%  \item FastMath, R/qtl, MINRES-QLP
%  \item OWL/RDF, AMR/multiphysics, Globus-Genomics
%\end{itemize}
%
